---
title: "electivity: Algorithms for electivity indices and measures of resource use versus availability"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

[![DOI](https://zenodo.org/badge/201148710.svg)](https://zenodo.org/badge/latestdoi/201148710)

Desi Quintans (2019). _electivity: Algorithms for Electivity Indices._ R package version 1.0.2. https://github.com/DesiQuintans/electivity

---

This package is essentially Lechowicz (1982) turned into an R package. It includes
all algorithms that were described therein plus the example data that
was provided for gypsy moth resource utilisation.

Lechowicz, M.J., 1982. The sampling characteristics of electivity indices.
Oecologia 52, 22–30. https://doi.org/10.1007/BF00349007

Users are encouraged to read the original paper before deciding which algorithm
is most useful for them. Lechowicz recommended Vanderploeg and Scavia's E* index
(implemented in this package as `vs_electivity()`) as "the single best, but not
perfect, electivity index" because "E* embodies a measure of the feeder's
perception of a food's value as a function of both its abundance and the
abundance of other food types present." In practice, he found that all indices
returned nearly identical rank orders of preferred hosts except for Strauss'
linear index (L).


## Installing

```{r echo=TRUE, eval=FALSE}
# Installing from CRAN (not yet!)
install.packages(electivity)

# Installing from GitHub
install.packages(remotes)
remotes::install_github("DesiQuintans/electivity")

library(electivity)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(electivity)
```


## Built-in datasets

```{r}
data(moth_distrib)  # Table 2 from Lechowicz (1982), raw data
data(moth_elect)    # Table 3 from Lechowicz (1982), calculated indices

head(moth_distrib)
head(moth_elect)
```


## Examples

### Calculating food preference for a single animal

```{r}
gypsy_moth_prefs <- vs_electivity(moth_distrib$r, moth_distrib$p)

names(gypsy_moth_prefs) <- moth_distrib$binomen

sort(gypsy_moth_prefs, decreasing = TRUE)
```

### Calculating food preference for many animals at many sites (Tidy approach)

```{r message=FALSE, warning=FALSE}
library(tidyr)
library(dplyr)
library(purrr)

# Example data
df <- tibble::tribble(
           ~snail,     ~site,      ~food, ~pieces_eaten, ~pieces_present,
           "Bert", "outside",   "Carrot",             3,               7,
           "Bert", "outside", "Broccoli",             3,               8,
           "Bert", "outside",     "Kale",             1,               1,
           "Bert",  "inside",   "Carrot",             5,              11,
           "Bert",  "inside", "Broccoli",             7,               3,
           "Bert",  "inside",     "Kale",             2,               4,
          "Ernie", "outside",   "Carrot",             6,               7,
          "Ernie", "outside", "Broccoli",             4,               8,
          "Ernie", "outside",     "Kale",             1,               1,
          "Ernie",  "inside",   "Carrot",             3,              11,
          "Ernie",  "inside", "Broccoli",             1,               3,
          "Ernie",  "inside",     "Kale",             4,               4
          )


prefs <- 
    df %>% 
    # Nest the data for each snail x site pair
    nest(-snail, -site) %>% 
    # Apply vs_electivity() (or any other function) to each nested dataframe
    mutate(score = map(data, ~ vs_electivity(.$pieces_eaten, .$pieces_present))) %>% 
    # Expand the result (score) into new rows
    unnest(score, data) %>% 
    # Omit unwanted columns
    select(-pieces_eaten, -pieces_present) %>% 
    # Turn the sites into columns that show electivity for each snail x food pair.
    spread(key = site, value = score)

knitr::kable(prefs, format = "markdown")
```